### 2012-11-24 MSWL_ECO_lesson1.tex
Initial version latex files

\chapter{Introduction and motivation}

\section{Introduction}

\section{Part I}\label{Part I}  % TODO: Change for properly section title.

\section{Part II}\label{Part II} % TODO: Change for properly section title.

...

\section{Conclusions}\label{conclusions}

### 2012-11-28 MSWL_ECO_lesson1.tex
#2 Adding introduction and motivation chapter. First class

@@
 
-\section{Introduction}
+\section{Introduction to Economic Aspects and FLOSS}\label{ECO_FLOSS} 
+The first assumption around FLOSS is how particular the FLOSS environment is.
+For this reason, business environment is also particular, and it has to be considered that means to make business have to be different from privative software business model, at least, at the beginning.\\
+\\
+\textbf{Privative software model}: It is based, basically, in two main activities:
+- Licenses: Based on payment around a license that allows software programs usage.Normally, code is protected.
+- Services: Support and other services around software. An activity that can be imitate on FLOSS software business model to make profit.\\
+\\
+This subject is focused, on the one hand, on how to make a profit business taking into account freedom of FLOSS.
+But, on the other hand, it focuses on other aspects, as, for example:\\ 
+\textbf{Is free software model "sustainable"?} For example: Will Libre Office competitive against Microsoft Office?\\
+\\
+Sustainability is very important for Free Software Projects to maintain alive in time, and this issue is not related to business around a particular FLOSS project.
+Both concepts are not exactly the same, although they are interrelated.
+If business models exists, sustainability over time is more probable\\
+\\
+There is an important issue to be resolved on facing business around Free Software: \textbf{identifying market niche}.
+In some particular cases, Free Software is not a good option, but the only option.
+Would Linux be today the same folowing a privative development model?\\
+\\
+Related to previous assumption, it has to be clarified that Free Software is not exactly altruism.
+Sometimes, Free Software is the best option in business.
+10 years ago, no considerations on Free Software was considered in companies.
+Nowadays, thorough considerations are taken, considering Free Software strategies on companies global business plans.\\
+\\
+Another Consideration will be carried out.
+It has to be differentiated which companies produce software against companies that are only Free Software consumers.
+Business models differ taking into account this issue as well.\\
+\\
+Other times, FLOSS is considered as the \textbf{best tool} to achieve a particular target.
+Free Software, for this reason,  is used as the best way to achieve a company goal.
+One example is Android Operating System.
+Android may not be generating big profits for Google, but Google, as company, accomplished the target to reach a position in the market that allows them not to depend on Apple or Microsoft Mobile Operating Systems.\\
+\\
+Another concept around Free Software is \textbf{Open Innovation}.
+This concept is very wide, and applies to innovation in very different fields, from Sports to Fashion.
+Open Innovation is interrelated to FLOSS, as it takes concepts and procedures from FLOSS projects environment.\\
+\\
+FLOSS Market is not an insignificant Market. To clarify some numbers around FLOSS economy, Foundation incomes can be studied.Some good examples around bids of Foundations:\\
+- Mozilla Foundation: around 100 Million Dollars\\
+- Open Stack: around 10 Million Dollars\\
+Apart from foundations, some enterprises are earning money.
+Red Hat, for example, is achieving profits in the last years.
+In 2011, total revenue was \$909.3 million, an increase of 22\% over the prior year, and subscription revenue was \$773.4 million, up 21\% year-over-year.
+\\
+\section{Market Numbers around FLOSS}\label{MARKET_FLOSS} 
+Some numbers can be analyzed around FLOSS.
+A good entry point is netcraft.
+This company analyzes Market based on Web Servers queries.
+They provide interesting reports, very trustable, as they are crawling machines to get the information, so the way the information is extracted is very different to more traditional reports based on surveys or other mechanisms.\\
+\\
+Related to Web Servers Market report analysis, it can be asserted that, in some cases, the market is leaded by a FLOSS Project (Apache in this case).
+Apache could get enough resources to keep being the leaders across the time.
+Moreover this, there are companies and individuals that make big effort on Apache.
+Companies as IBM, for example, did not try to create a web server, but instead they bet for Apache, as they do not want its competitors to gain the market.
+Apache has reached to be a standard de facto, so to sell a Web Server different from Apache, you can not sell anything the same or with less performance than Apache, especially when it is free.\\
+\\
+Around Browsers Market, however, other curious paradigm is happening: Webkit is base for Safari and Android Browser, and both are the leaders for Mobile Browsers.
+Curious as Google and Apple fight each other in many other businesses.\\
+\\
+Analyzing Market Reports some basic concepts can be foregrounded: 
+- \textbf{Unstable Market}: leader having less than 60\% of the Market (Web Browsers Market, for example is an unstable Market).\\
+\\
+On the other hand, some other considerations arise around particular movements made by Companies around Software business, and how they change Market.
+Apple, for example,  hanged the Market when it allowed application to be sold with no exclusive rights, but with a percentage (30\%, a big amount taking into account the number of applications sold).
+On the other hand, Google introduced a wider and opener strategy, with a Free Operating System too.
+They are both examples on how selection of FLOSS strategy is important, as nobody thought Android could fight against Apple.\\
+\\
+Another interesting report shows that, in year 2008, many enterprises (around 40\%) were not using free software, but were interested on it.
+Other 30-35\% were not interested.
+Nowadays, everybody having a big/medium company knows that FLOSS is key business for their IT models.\\
+\\
+Another report related to Big computers market was shown.
+The main question around Super Computer Market, nowadays, is not who uses Linux, but wich distribution of Linux use each Super Computer.
+Linux Kernel is the undisputed leader in the market and, however, it is also a key player in many other different markets, as tMobile Devices.
+That is really something new in informatics, and has to do with openness, adaptability, etc.
+It is curious that, in Desktop, Linux never ends up entering the Market, when it is an example of a nice Operating System running on very different systems.\\
+\\
+To summarize, it has to be stated that FLOSS is a particular case, in technology and, of course, in economics.\\\\
+It also has to be considered that, Internet is changing the model.
+First in time, things were done by small groups in small towns.
+Later, things were done by Big Enterprises or States.
+Nowadays, we are goint back to the "small town times", but this time small towns are interconected, as Internet is used to synchronise people.
+It also has to be clarified that,  sometimes, towns are not so small, as in the Wikipedia case.\\
+\\
+Economical traditional theories count on consumers on producers, and resources that one generates, and others consume.
+Market Theory, Supply and Demand, Value Chain, etc.
+In FLOSS somethings happen in the same format.
+For now Linux to be widely used, Linux had to be started 20 years ago.
+The difference is that traditional markets rules are not appliable:\\
+- First of all, there is no price. With no price, 90\% of theories are not valid. 
+- Apart from that, a company paying money to a FLOSS project (i.e.:Mozilla Foundation) can not valuate how much money will recover or win, at least directly.\\ 
+- But another question arises:In fact, is it relevant?
+Is relevant to know how much money will a Company pay to acquire Mozilla?
+Market theories can not explain all the models and cooperation ways.
+There are economist that study and specialized in this paradigm.\\
+\\
 
-\section{Part I}\label{Part I}  % TODO: Change for properly section title.
+\section{A new paradigm in internet}\label{MARKET_FLOSS}
+
+A video was played in class to demonstrate the new paradigm around Internet. The most important conclussions after watching are:\\
+1) Crowdsourcing: Multiple people doing something. New economic value.\\ 
+2) It is difficult to start some kind of projects (Radio, Television, Newspaper, etc.).\\ 
+3) In other similar markets (super-computer), some time (2002), the biggest supercomputer was SETTI program (were thousands of contributors supplied his limited resources).\\
+4) Things are changing: Wikipedia, Apache, are FLOSS success stories.\\
+5) Crowdsourcing (SETI): it is something that can be very huge, involving many people, and done via Internet, by means different from states or big companies.\\
+6) Computer Market is high advanced, Computers are cheap and gives you freedom to:\\
+- Do whatever you want with it.\\
+- Share it to do other things, help others.\\
+7) The most collaborative company might be Google.
+Google page rank algorithm takes into account what the rest of the people consumes, and what the rest of the people link.
+They won Open-Directory (Free Software).
+They take profit of what people does not value, or what people want to do.\\
+There are cases as Google Goggles or Scanning words projects in Captchas.\\
+\\
+\begin{tabular}{| l | c | r |}
+   \hline
+   & Market Based & Non-Market Based \\
+   \hline
+   Decentralized  & Price Based    & \textbf{Social Sharing/Exchange} \\
+   \hline
+   Centralized    & Firm Hierarchy & Government/non profits (NGOs) \\
+   \hline
+\end{tabular}
+\\
+\\
+(*1)In past time this paradigm was used for small projects (look after of a garden in a neighbourhood)\\
+It is normally valid for "non-material" things, changing the sense of property. That happens on software, music, etc.\\
+Non-market decentralized models need new means of coordination.\\
+Means competition: (FLOSS vs Microsoft, Skype vs Telecomms, Wikipedia vs Encyclopedias, etc.)
+The results of the competition means also the kind of the society we will live.
 
@@
 
-...
-
 \section{Conclusions}\label{conclusions}

### 2012-12-03 MSWL_ECO_lesson2.tex
Initial version latex files
fixed #4 - Summary: Impact on stakeholders and markets, strategic aspects
fixes #4: Revision: Impact on stakeholders and markets, strategic aspects

--- 2012-11-24 MSWL_ECO_lesson1.tex
@@
-\chapter{Introduction and motivation}
+\chapter{Impact on stakeholders and markets, strategic aspects}
 
-\section{Introduction}
+This section starts by analysing an hipotetical situation.
+Let's supose that we have been hired as by a company with certain business, and we are taken into account as decision making advisors in the area of technology.
+The company has already made a decision about releasing their key software product under Apache or GPL license, and there is a new meeting to decide which license to choose?
+And more specifically: \textbf{Which license will attract more companies to collaborate in our business?}\\
 
-\section{Part I}\label{Part I}  % TODO: Change for properly section title.
+We recall that the Apache license type can be classified as permissive, while a GPL as copyleft, each with its own implications. \\
 
-\section{Part II}\label{Part II} % TODO: Change for properly section title.
+The debate at class started with different approaches, some were inclined to the GPL because it ensured the continuity of the project under the same trend, others by Apache because commercial companies in general have less resistance to such licenses, and others ultimately could not come to any conclusion.
+Eventually all converged to that there is no formula with calculated results, the choice of a FLOSS license type usually depends on at least the following premises:
 
-...
+\begin{itemize}
+\item Desired goal to be obtained with the product
+\item Niche market to which the product is intended
+\item Presence of the company in the current market
+\item Preliminary study of the current market and definition of the possible innovation
+\item Among many others
+\end{itemize}
+
+Overall we can see that the decision making regarding to Free Software, is not far from other business areas and corresponding analysis, every FLOSS product deserves its specific considerations.
+
+\section{Impact on companies using software and in the production fabric}
+
+One thing is very clear when talking about the economic impact that a business strategy can produce, and it is to be focused from a purely practical point of view.
+In this sence we watched the video\footnote{http://www.parc.com/event/1092/open-for-business.html} of a lecture at PARC (Palo Alto Research Center) about \textquotedblleft Open for business: Building successful commerce around open source", whose presenter was M\aa rten Mickos, Eucalyptus Systems CEO and ex-CEO of MySQL AB, two successful OSS companies.\\
+
+Some main ideas about the presentation are:
+
+\begin{itemize}
+\item A business presentation, not a technical one
+\item Make a lot of money with open source is the main goal, because companies are the ones who finances developers and so on, if it requires closed source features or going to the enterprise market, it has to be done in order to sustain
+\item For many people is tough talking about FLOSS and business at the same time, because there are mistaken concepts around these topics together
+\item FLOSS a better method to produce software (goods), faster (communities), common interest into solve problems (developers), relies over internet (distribution), users are happy with it (customers), but difficult to sell it (companies)
+\item When talking from a practical point of view, it important to differentiate FLOSS from philosophical issues involved
+\item Forms of production and distribution are very important, and in this case on intangible assets
+\item Maybe FLOSS works because there are people trying to solve their own personal problems over this model, so maybe there is no charity or voluntary at all
+\item Important companies such as RedHat or Oracle have inverted lots of money on FLOSS and have shown it is a sustainable model
+\item Now when talking about FLOSS, words like: serious, commercial and professional, show up
+\item Some time ago vendors didn't care about FLOSS, but now customers are demanding it
+\item Before quality of software was dictated only by proprietary software companies, but now with FLOSS the quality is in sight of all people
+\item The core-business for a company, may be non-core for others, this is a good feature to take advantage of
+\item Eventually a company that successfully penetrates with FLOSS a market that was proprietary, tends to enforce the rules
+\item FLOSS success vs Commercial success graphic:
+\begin{figure}[h]
+\begin{center}
+\includegraphics{success}
+\caption{FLOSS success vs Commercial success}
+\label{fig:success}
+\end{center}
+\end{figure}
+\end{itemize}
+
+\section{The Economic Motivation of Open Source Software: Stakeholder Perspectives}\label{Part II}
+
+In working process... by Daniel G\'amez.\\
 

### 2012-12-14 MSWL_ECO_lesson2.tex
fixes #5: Revision: Impact on stakeholders and markets, strategic aspects

@@
 \item FLOSS a better method to produce software (goods), faster (communities), common interest into solve problems (developers), relies over internet (distribution), users are happy with it (customers), but difficult to sell it (companies)
-\item When talking from a practical point of view, it important to differentiate FLOSS from philosophical issues involved
+\item When talking from a practical point of view, it is important to differentiate FLOSS from philosophical issues involved
 \item Forms of production and distribution are very important, and in this case on intangible assets
@@
 \item Important companies such as RedHat or Oracle have inverted lots of money on FLOSS and have shown it is a sustainable model
-\item Now when talking about FLOSS, words like: serious, commercial and professional, show up
+\item Now when talking about FLOSS, show up words like: serious, commercial and professional
 \item Some time ago vendors didn't care about FLOSS, but now customers are demanding it
@@
 \item Eventually a company that successfully penetrates with FLOSS a market that was proprietary, tends to enforce the rules
-\item FLOSS success vs Commercial success graphic:
+\item FLOSS success vs Commercial success graph, shows the opposite limits that have reached eg.
+Apache and Microsoft products, as well as products like MySQL DB or Linux kernel occupying the midpoint:
 \begin{figure}[h]
@@
 \includegraphics{success}
-\caption{FLOSS success vs Commercial success}
+\caption{FLOSS success vs Commercial success.}
 \label{fig:success}
@@
 
-In working process... by Daniel G\'amez.\\
+An important consequence that FLOSS brings to the economic market is the emergence of new players in the system.
+Now there is not only a software manufacturer, channels and customers, but there are also private developers, system integrators and communities that have as much presence as any powerful company in the market of a certain field.\\
+
+From system integrators' perspective, we interpretated a theorical \textquotedblleft IT solutions demand curve"\footnote{http://dirkriehle.com/computer-science/research/2007/computer-2007-article.html}, in which highlights three main factors of IT: hardware, software, and services.\\
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics{curve}
+\caption{Sales margins and number of customers.
+(a) The lower price limit determines the customers the system integrator takes on, (b) switching from closed source software to open source software can result in more customers and higher profits.}
+\label{fig:curve}
+\end{center}
+\end{figure}
+
+Based on the fact that the hardware is usually maintained at a stable price, by reducing the costs associated with licensing fees when using FLOSS software, integrators can vary as desired, whether the services provided or the price associated with such services.
+This represents a great advantage over the use of proprietary software, in the sense that profits areas are more flexible, that is, starting from a certain cost, theoretically can be increased the number of services offered, or could be decreased the price of services in order to attract more customers.\\
+
+Also, depending on whether the product target is massive or not, will require the allocation of a greater or lesser value to recover the investment later.
+Some examples were mentioned, as for instance:
+
+\begin{itemize}
+\item The market for mobile devices, led in part by Google nowadays, is a consequence of the low cost of the software, allowing direct investment almost only into manufacturing hardware, and therefore having a competitive advantage in the price at which the product is offered
+\item Android Operating System on any of the tablets from Barnes \& Noble or Samsung, versus Apple's
+\end{itemize}
+
+Another concept handled was the cost and the price of intangible assets, who behave differently from tangible ones.
+From the closed source software vendor perspective a leading software vendor sets a price that maximizes its profits, offering the product to different customers at the same price.
+Usually most of the investment comes from shipping the first copy of the software.
+On the other hand in an open source community, given the right license, anyone can set up a company and start selling the software provision, maintenance, and support.
+Pricing will be based on markup over cost, with a high markup new companies will enter the market, with a low markup companies will leave the market.
+The first copy of the software is produced in a community without losing the appeal of the product, not necessarily in a cheaper way.
+Customers benefit from this situation, because they get a much cheaper price compared to proprietary software, and FLOSS vendors also benefit because they displace proprietary products.\\
+
+There is a group of graphics that show a long-run average cost curve for a single software product\footnote{http://dirkriehle.com/computer-science/research/2007/computer-2007-article.html}.
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics{cost-price}
+\caption{Differences in cost and pricing.
+(a) and (c) show a similar average cost per unit for closed source and open source software, while (b) and (d) indicate the relationship between price and number of units sold for open and closed software.}
+\label{fig:cost-price}
+\end{center}
+\end{figure}
 
 \section{Conclusions}\label{conclusions}
+
+In working process... by Daniel G\'amez.\\

### 2012-12-17 MSWL_ECO_lesson3.tex
Initial version latex files
fixes #6 - Summary: Cost/benefit analysis. Cost/benefit analysis section.

--- 2012-11-24 MSWL_ECO_lesson1.tex
@@
-\chapter{Introduction and motivation}
+\chapter{Cost/benefit analysis}
 
@@
 
-\section{Part I}\label{Part I}  % TODO: Change for properly section title.
+\emph{Why a company or public institution would base its strategy on FLOSS ?} With this initial question we led the argument that a company would gain by associating their business model to FLOSS projects
 
-\section{Part II}\label{Part II} % TODO: Change for properly section title.
+Most companies are interested in products not if its FLOSS or not.
+A company wants the service and doesn't care about its origin, in most cases.
+Wants the fiability and consistence of the product that are buying.
+At the other hand we have integration companies that are near to FLOSS products development that allows them a deeper analysis capabilities.
 
-...
+\subsection{Cost/benefit analysis}
+
+We present an analysis of costs / benefits related to FLOSS products shown on consistency beyond the brand name starting with one sentence: \emph{FLOSS projects are available to anyone with a computer and an internet connection}.
+
+\begin{enumerate}
+
+    \item Advantages of public scrutiny and chances for improvements
+    \begin{itemize}
+        \item Compare benchmarks between projects and its goals.
+Could analyse every code line as you want and check every feature that has the product and compare with other projects using standar quality models\footnote{QSOS, OpenBRR, QUALOSS} adapted to your interests.
+And of course, anyone can view the source code of your project so you have to get it right and well documentated.
+    \end{itemize}
+
+    \item Real competition in development and maintenance
+    \begin{itemize}
+        \item In private projects there is only a main company that could develop improvements in software.
+Maintenance is very important and every new business model could grow around FLOSS projects.
+So your product as a costumer depends only of one company, what would happen if the company goes bankrupt?
+    \end{itemize}
+
+    \item Technical feasibility vs. marketing
+    \begin{itemize}
+        \item Its technical feasibility is its own marketing.
+Private Software only has its own reports about its software, so you only could trust them without any test.
+In some cases, allowed some evidence but you can not be sure that what you are testing is the final product so you are limited by what the company offers.
+Choose before (private) or after (free) the trial.
+    \end{itemize}
+
+    \item Easy testing, deployment, update
+    \begin{itemize}
+        \item Easy update to continue retrieving improvements.
+If the product is FLOSS you can update, test and deploy yourself the product because you have all needed tools, source code.
+Of course you can pay another company for develop updates in the product too and continue focused in your issues.
+    \end{itemize}
+
+    \item Lower barriers to business opportunities
+    \begin{itemize}
+        \item Infraestructure around FLOSS is cheaper and lets you focus in your own business.
+Lets you focus only in your business niche.
+If you need to add something that could be interesting to your model you can include your goals in FLOSS Roadmaps, it's not easy but it's possible despite in the private roadmaps solutions.
+FLOSS products gives you easy technology access.
+    \end{itemize}
+
+    \item Non-formal collaboration and pooling of resources
+    \begin{itemize}
+        \item FLOSS projects could have competitors working together in the project.
+As an example we have AMD and Intel working together in Linux even Microsoft for first time\footnote{http://www.linuxfoundation.org/news-media/announcements/2012/04/linux-foundation-releases-annual-linux-development-report}.
+Google and Apple contribute with Webkit as main companies\footnote{http://trac.webkit.org/wiki/WebKit\%20Team}.
+Because of community rules and make a common infraestructure for all.
+NDA doesn't exist.
+    \end{itemize}
+
+    \item Technology transfer
+    \begin{itemize}
+        \item Low barriers to share knowledgement.
+Web is used as a tool thanks to FLOSS communities.
+\emph{if each one has an apple and we exchange apples we still have each but if we each have an idea and we exchange, the two will have two ideas.}
+    \end{itemize}
+
+    \item Libre software as an strategic tool
+    \begin{itemize}
+        \item FLOSS as part of your strategy.
+Use FLOSS products to reach to your desired business niche, as a tool to easily reach your goals.
+We have the sample of Google that wants advertisement benefits in mobiles and used FLOSS (Android) to achieve its goal in mobile market to make people use their services easily.
+    \end{itemize}
+
+\end{enumerate}
+
+\subsection{What disappears when a business uses libre software?}
+
+\textbf{TBC}
 
 \section{Conclusions}\label{conclusions}
+
+\textbf{TBC}

### 2012-12-18 MSWL_ECO_lesson2.tex
fixes #4: Revision: Impact on stakeholders and markets, strategic aspects

@@
-\chapter{Impact on stakeholders and markets, strategic aspects}
+\chapter{Impact on stakeholders and markets, strategic aspects}\label{Chapter 2}
 
@@
 
-The debate at class started with different approaches, some were inclined to the GPL because it ensured the continuity of the project under the same trend, others by Apache because commercial companies in general have less resistance to such licenses, and others ultimately could not come to any conclusion.
+The debate at class started with different approaches, some were inclined to the GPL because it ensures the continuity of the project under the same trend, others by Apache because commercial companies in general have less resistance to such licenses, and others ultimately could not come to any conclusion.
 Eventually all converged to that there is no formula with calculated results, the choice of a FLOSS license type usually depends on at least the following premises:
@@
 
-\section{Impact on companies using software and in the production fabric}
+\section{Impact on companies using software and in the production fabric}\label{Section 2.1}
 
 One thing is very clear when talking about the economic impact that a business strategy can produce, and it is to be focused from a purely practical point of view.
-In this sence we watched the video\footnote{http://www.parc.com/event/1092/open-for-business.html} of a lecture at PARC (Palo Alto Research Center) about \textquotedblleft Open for business: Building successful commerce around open source", whose presenter was M\aa rten Mickos, Eucalyptus Systems CEO and ex-CEO of MySQL AB, two successful OSS companies.\\
-
+In this sence we watched the video\footnote{http://www.parc.com/event/1092/open-for-business.html} of a lecture at PARC (Palo Alto Research Center) about \textquotedblleft Open for business: Building successful commerce around open source", whose presenter was M\aa rten Mickos, Eucalyptus Systems CEO and ex-CEO of MySQL AB, two successful OSS companies.
 Some main ideas about the presentation are:
@@
 
-\section{The Economic Motivation of Open Source Software: Stakeholder Perspectives}\label{Part II}
+\section{The Economic Motivation of Open Source Software: Stakeholder Perspectives}\label{Section 2.2}
 
@@
 
-Based on the fact that the hardware is usually maintained at a stable price, by reducing the costs associated with licensing fees when using FLOSS software, integrators can vary as desired, whether the services provided or the price associated with such services.
+Based on the fact that the hardware is usually maintained at a stable price, by reducing the costs associated with licensing fees when using FLOSS integrators can vary as desired, whether the services provided or the price associated with such services.
 This represents a great advantage over the use of proprietary software, in the sense that profits areas are more flexible, that is, starting from a certain cost, theoretically can be increased the number of services offered, or could be decreased the price of services in order to attract more customers.\\
+
+\pagebreak
 
@@
 
-\section{Conclusions}\label{conclusions}
+\section{Conclusions}\label{Conclusions Chapter 2}
 
-In working process... by Daniel G\'amez.\\
+Making a general analysis we can see that through FLOSS there is a viable business option, ways to compete in the market and new business opportunities.
+It is important to make a preliminary assessment of the business to determine whether or not the FLOSS fits on it.
+In many cases FLOSS is the most efficient, but not always, anyway FLOSS is highly adaptable and conformance to needs, it will always provide an alternative to which pay attention at.\\
+
+It has been said also that this is a service economy, bringing software as a service, instead of software as a product.
+This changes traditional rules and allows opportunities for new stakeholders on certain fields, with the premise of innovate to get real competitiveness on the market.\\
+
+The benefits provided by the FLOSS are endless, but it is necessary to understand them, understand their environment, assume the appropriate role, be innovative and make the best of it, as in any business strategy.

### 2012-12-19 MSWL_ECO_lesson3.tex
updated #6 - Summary: Cost/benefit analysis
* Updated section "What disappears when a business uses libre software?"
* gitignore file avoiding eclipse texlipse project files.
* main file updated removing conflict

@@
 
-\subsection{What disappears when a business uses libre software?}
+\section{What disappears when a business uses libre software?}
+
+What brings us to a business using FLOSS technologies? What doors are open to us?
+
+\begin{itemize}
+    \item \emph{Dependence on monopolies:}
+    \begin{itemize}
+        \item Real competition, better products, better services.
+The technology is not dependent on a single supplier, technology is the union of different suppliers in oneself when using the product, it can become a provider of the same technology.
+    \end{itemize}  
+    \item \emph{Importance of vendor reliability:}
+    \begin{itemize}
+        \item Future depends on product acceptance in the market.
+The vendor is exposed to continuous product analysis by different users.
+In every field of use of the product are certain tests that help the overall assessment and therefore the acceptance thereof by users.
+This process causes the provider to increase its reliability with regard to the consumer.
+    \end{itemize}  
+    \item \emph{Decisions taken based on few elements:}
+    \begin{itemize}
+        \item Software can be tested in its real environment, at a very low cost.
+You have the possibility to imagine all kind of test in every environment you choose to check product capabilities and compare your results with other analysis.
+    \end{itemize}  
+    \item \emph{Dependence on the strategy of providers:}
+    \begin{itemize}
+        \item Decisions on the evolution of a product taken for those contributing with resources.
+You have the possibility to include your thought into the product roadmap, also you can improve with your contributions to the project and share ideas with the community in an cross-ideas open platform.
+    \end{itemize}  
+    \item \emph{Confidence on "black boxes":}
+    \begin{itemize}
+        \item Using FLOSS projects we can see the whole process of the data inside de technology, we don't have \emph{black holes} where we miss which operations are done.
+We do not lose the path of process operations being applied to our data the software from input to output.
+For example an application service like gmail.
+This application hides we treat our data to drive the process of sending an email and the result shows that message has been sent as output
+    \end{itemize}  
+    \begin{figure}[h]
+        \begin{center}
+            \includegraphics{blackbox}
+            \caption{BlackBox; we only can see the input and the output, not the process.}
+            \label{fig:curve}
+        \end{center}
+    \end{figure}
+\end{itemize}
+
+\section{Conclusions}\label{lesson-3-conclusions}
 
 \textbf{TBC}
-
-\section{Conclusions}\label{conclusions}
-
-\textbf{TBC}

### 2012-12-19 MSWL_ECO_lesson4.tex
Initial version latex files
updated #8 - Summary: Business models
- First part of the lesson in Spanish prepared for translation into
English.

--- 2012-11-24 MSWL_ECO_lesson1.tex
@@
-\chapter{Introduction and motivation}
+\chapter{FLOSS business models}
 
-\section{Introduction}
+\section{Introduction}\label{lesson-4-introduction}
 
-\section{Part I}\label{Part I}  % TODO: Change for properly section title.
+FLOSS
+
+The one million dollar question \emph{Which model has success.}
+
+Different FLOSS business models inside the market. Those sample are from two years ago, so don't take any strict conclusions because this models are continusly evolving.
+
+\subsection{Origin}\label{lesson-4-origin}
+
+Open innovation rise business models.
+
+Activos de la empresa para ganar dinero, cultura, empleados y conocimiento a parte del capital monetario que dejaremos de lado. Open innovation stares at knowlegment and its use.
+
+Empezaremos con el ejemplo de una empresa donde su producto es el referente y no el conocimiento de los empleados para fabricar el producto, el caso sería el siguiente:
+
+Si sólo se produce el producto estrictamente como dicta la empresa aunque exista un empleado que sepa construir el producto mucho mejor de lo que está haciendo la empresa por lo tanto está infrautilizado, el recurso se aprovecha al máximo.
+Al contrario que unificar el conocimiento entre los empleados para consesuar una mejor forma de construir el producto en la que participan todos y las mejoras o sugerencias son escuchadas tendiendo a evolucionar ya que nada es permanente.
+
+¿ Como gestionan las empresas el conocimiento ?
+\emph{El conocimiento es básico para la empresa y por ello ha de protegerlo}, esto rompe con los pricipios del FLOSS, para que la competencia no compita con su propio conocimiento.
+
+El tratamiento privativo del conocimiento restringe el radio de mejora frente a los diferentes casos comunes en los que se comparte el conocimiento, ya que de esta forma al compartir el conocimiento ambos ganan.
+La innovación rápida frente a la protección de la empresa del entorno en donde se cierra a opiniones externas.
+
+El intercamio de conocimientos comienza entre el intercambio de empleados por parte de las empresas, como ejemplo el caso de Silicon Valley a mediados de la década de los 70 en donde las empresas necesitaban ingenieros y la vida de un ingeniero en una empresa era de 18 meses por lo que fluía el conocimiento entre todas las empresas hasta concevir una masas crítica a partir del intercambio de conocimientos por el trasvase de empleados.
+
+\begin{itemize}
+  \item Case Study: \emph{Gold Corp} Modelo predictivo.
+  \item Case Study: \emph{Netflix}
+\end{itemize}
+
+Modelos predictivos desarrollados por personas externas a la misma empresa premiando al modelo ganador.
+
+Empezando por los NDAs (non-disclosure agreement) como puntal en las empresas de Software, acotan la comunicación y la difusión del conocimiento.
+Hasta que vino la brecha en donde la gente creyó en que la innovación no debería estar sujeta únicamente a la empresa y que de esta manera crecería más rápidamente el mismo, la realimentación y el contacto con 'la competencia'.
+Este modelo es extrapolable a cualquier otro negocio, como ejemplo la inclusión de los clientes en el proceso de diseño recibiendo un feedback de la comunidad de usuarios.
+
+Kickstarter: dejando en manos de los clients la prospectiva de mercado previamente a que el producto se fabrique.
+
+\section{Current trends}\label{lesson-4-current-trends}
+
+\emph{"There is no silver bullet"}
+
+Estar atento al mercado, no sólo al que se pertenece si no a los mercados externos por lo que se podría definir como una especialización del mercado o expertización.
+Se premia la colaboración con los usuarios generando conocimiento a la par que la innovación, por lo que el usuario que tenga interés puede mejorar el producto.
+
+\section{False myths}\label{lesson-4-false-myths}
+
+El precio de venta es el que está dispuesto a pagar la gente, no tiene nada que ver con el coste de producción.
+
+El valor que le da el usuario/cliente puede evaluarse:
+
+\begin{itemize}
+    \item El valor del uso de la herramienta, inversión. Ahorro de 5000 euros por un ordenador pero se paga 500 euros por el como valor de venta.
+    \item El valor de venta.
+\end{itemize}
+
+El 75\% de los programas se dedica al mantenimiento mientras que el resto se dedica a innovación.
+
+El valor del software no se mide a partir del \emph{valor de reemplazo}.
+No importan los costes de producción si no de cuánto costaría tener que producir el producto de nuevo, asociado a los bienes físicos.
+
+Expected value of the future service.
+
+EL límite alto en el software está marcado por el servicio futuro que me ofrece, mirando hacia adelante no lo que ha hecho hasta ese momento.
+Estabilidad del software, soporte, documentación, adaptación o como se puede conseguir una certificación.
+A diferencia de los bienes físicos no se piensa en cuanto se puede conseguir el mismo producto actualizado mientras que en un físico si que se mira atrás.
+
+    Services == Source of revenues
+
+En Software no compras el producto, valoras el servicio que te ofrece.
+
+La organizaciónd el modelo de Software Libre permite ofrecer mejores servicios:
+\begin{itemize}
+    \item Usuarios que se convierten en clientes.
+    \item escalabilidad en identificación de fallos. Los programas de éxito suelen tener menos errores, el éxito medido en usuario.
+    \item Compartir riesgos y costes de producción. Entre los actores se comparte el gasto del desarrollo.
+    \item Prácticas monopolísticas muy difíciles. Un fork permite escapar de la monopolización del producto.
+\end{itemize}
+
+\emph{Si no tienes usuarios no tienes modelo de negocio.}
 

### 2012-12-27 MSWL_ECO_lesson4.tex
updated #6 - Summary: Cost/benefit analysis
Comment code in spanish to avoid compiling error using verbatim package.
Add my name to the authors and translate from spanish to english.

@@
 Open innovation rise business models.
+
+% I've commented this section to avoid compiling problems. commet is defined in
+% verbarim package, use it as a normal item after importing in main file.
+
+\begin{comment}
 
@@
 
-En Software no compras el producto, valoras el servicio que te ofrece.
+In Software you don't purchase the product , you evaluate the service that product offers.
 
-La organizaciónd el modelo de Software Libre permite ofrecer mejores servicios:
+The organization of Free Software model allows offering best services:
 \begin{itemize}
-    \item Usuarios que se convierten en clientes.
-    \item escalabilidad en identificación de fallos. Los programas de éxito suelen tener menos errores, el éxito medido en usuario.
-    \item Compartir riesgos y costes de producción. Entre los actores se comparte el gasto del desarrollo.
-    \item Prácticas monopolísticas muy difíciles. Un fork permite escapar de la monopolización del producto.
+    \item Users who become clients.
+    \item Scalability in failure identification. Successful programs use to have less errors, where success measured in user.
+    \item Share risks and production costs. Development cost is shared among the actors.
+    \item Monopolistic practices are very difficult. A fork allows to escape from product monopolization.
 \end{itemize}
 
-\emph{Si no tienes usuarios no tienes modelo de negocio.}
+\emph{If you don't have users , you don't have any available business model.}
+\end{comment}
 

### 2012-12-28 MSWL_ECO_lesson4.tex
Finish translation from Spanish to English

--- 2012-11-24 MSWL_ECO_lesson1.tex
@@
-\chapter{Introduction and motivation}
+\chapter{FLOSS business models}
 
-\section{Introduction}
+\section{Introduction}\label{lesson-4-introduction}
 
-\section{Part I}\label{Part I}  % TODO: Change for properly section title.
+FLOSS
+
+The one million dollar question \emph{Which model has success.}
+
+Different FLOSS business models inside the market. Those sample are from two years ago, so don't take any strict conclusions because this models are continusly evolving.
+
+\subsection{Origin}\label{lesson-4-origin}
+
+Open innovation rise business models.
+
+ I've commented this section to avoid compiling problems. commet is defined in
+ verbarim package, use it as a normal item after importing in main file.
+
+%\begin{comment}
+\subparagraph{Main Content}
+
+Assets of the company to make money, culture, employees and knowledge as a part
+of the capital that we will put aside money.
+
+Open innovation stares at knowlegment and its use.
+
+We begin with the example of a company where its product is the reference and
+not the knowledge of its employees to manufacture the product, this case would
+be:
+
+If only produces the product strictly like a dictate of the company , although
+exists an employee who knows how to manufacture the product more better than the
+company its self therefore is underused, the resource is maximized.
+
+Unlike the knowledge unification among employees so as to find the best way for
+developing the product in which they all participate and the best improvments or
+suggestions tending to envolve are welcomed as nothing is permanent.
+
+How companies manage the knowledge ? \emph{The knowledge is essential for the
+company and therefore it has to protect}, this is against with the
+principles of  FLOSS, so as the competion does not compete with their
+knowledge. The private treatment of the knowledge restricts the radius of
+improvement over the different cases in which the knowledge is being shared because by sharing knowledge they both
+win. The rapid innovation against the protection of the company - of the
+enviroment where closes to outside opinions.
+
+The knowledge exchange begins between employees by companies
+as for example in case of Silicon Valley at the middle of 70's where the
+companies needed engineers and an engineer's life in a company was no more than
+18 months so the knowledge was flowing between all companies until convincing a critical mass
+knowledge exchange for the transfer of employees.
+
+\begin{itemize}
+  \item Case Study: \emph{Gold Corp} Predictive model.
+  \item Case Study: \emph{Netflix}
+\end{itemize}
+
+Predictive models developed by people outside of the company rewarding the ``winning model''
+
+Beginning with NDAs (non-disclosure agreement) as a point in Software
+companies,they delimit communication and dissemination of knowledge.
+
+Until the gap came in where people believed that innovation should not be the
+solely subject to the company and that in this way it would grow faster, the
+feedback and the contact with 'competition'.
+
+This model could be extrapolated to any other business, for example the
+inclusion of customer in the design process to receive feedback from the user community.
+
+Kickstarter:Leaving the market foresight to the customers before the prouduct is
+being manufactured
+
+\section{Current trends}\label{lesson-4-current-trends}
+
+\emph{"There is no silver bullet"}
+
+Be attentive to the market , not only to which one belongs but to external
+markets so that can be defined as a specialization or an expertizing of the
+market.It rewards collaboration with users generating knowledge on par with innovation
+so that the interested user could improve the product.
+
+\section{False myths}\label{lesson-4-false-myths}
+
+The sales price is the price which people are able to pay , and it has nothing
+to do with the production cost. 
+The value that is given by user/client can be evaluated :
+
+\begin{itemize}
+    \item The value of using the tool , investment.
+    Savings of 5000 euros  for one computer but you pay 500 Euros as a sale
+    value.
+    \item The sales value.
+\end{itemize}
+
+The 75\% of the programs is dedicated to the maintenance while the rest is
+dedicated to the innovation.
+The value of software is not measured by the \emph{value of replacement}.
+
+It doesn't mind the cost of the production but what it would cost if you don't
+having to produce the product again, associated with the physical assets.
+
+Expected value of the future service.
+
+The upper limit in the software is marked by future service that offers to me ,
+by looking forward in what he has not been done so far.
+
+Software stability , support, documentation , adaption or how can you get a
+certification.Unlike the physical goods you don't think as soon as you can find
+the same product as they can get the same product updated while that in a
+physical if you look back.
+
+    Services == Source of revenues
+
+In Software you don't purchase the product , you evaluate the service that product offers.
+
+The organization of Free Software model allows offering best services:
+\begin{itemize}
+    \item Users who become clients.
+    \item Scalability in failure identification. Successful programs use to have less errors, where success measured in user.
+    \item Share risks and production costs. Development cost is shared among the actors.
+    \item Monopolistic practices are very difficult. A fork allows to escape from product monopolization.
+\end{itemize}
+
+\emph{If you don't have users , you don't have any available business model.}
+%\end{Main Content}
 

### 2013-01-15 MSWL_ECO_lesson5.tex
Initial version latex files
Update MSWL_ECO_lesson5.tex
My dear classmates, this my first Latex and first time i deal with github  please let me now if there is something wrong :)

--- 2012-11-24 MSWL_ECO_lesson1.tex
@@
-\chapter{Introduction and motivation}
+\chapter{Libre software and open innovation}
 
@@
 
-...
+% open innovation and open source .
+%Patterns of open innovation in free software
+% done by Amal Roumi
+
+\section* {Open source and Open innovation} \label{Open source and Open innovation}
+
+“Open innovation is a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology”
+‘Open innovation’ is a term coined by Professor of Business Henry Chesbrough in his 2003 book Open Innovation: The New Imperative for Creating and Profiting from Technology.
+
+In the years since its publication, Chesbrough’s ideas on how technology should be managed and exploited have become extremely influential.
+Over the same period, the public profile of free and open source software (FOSS) has risen.
+So we can say that FOSS is an example of open innovation in software.
+They have two main element: use the technology, and collaborative development of that technology, unlike many individual participants, companies must also consider an economic return to justify there investment in.
+
+\subsection* {Patterns of open innovation in free software} \label{Patterns of open innovation in free software}
+There are four patterns of open innovation in free software:
+\begin{itemize}
+ \item  \textbf {Pooled RnD}: Pool resources to innovate in a common platform, exploit results.
+  \begin{itemize}
+		\item Maximization: concentrate in their own needs.
+		\item Incorporation: shared technology in their products.
+		\item Motivation: pool of contributors assumed.
+	\end{itemize}
+Lunix, Mozilla For both donate RnD to open source project while exploiting the pooles RnD of all contributors to facilitate the sale of related products.
+
+ \item  \textbf {Spintouts}: Release the poetical of technologies within the firm that are not creating value.and they have opportunities to release more value from their technologies by position them out the firm but at the same time maintaing an ongoing corporate involvement.
+Here the “spinout”come.
+	\begin{itemize}
+		\item Maximization of impact of non-core technologies.
+		\item Incorporation of contributions by third parties.
+		\item Motivation: self-sustainable (or less resource-consuming) communities.
+	\end{itemize}
+Eclipse, Beehive,Jikes projects the sponsor firms spinout open source project that were closely aligned with firm’s ongoing strategies
+
+ \item \textbf{Selling complements:} Income from complements, shared innovation in a common core.
+	\begin{itemize}
+		\item Maximization by centering on core products.
+		\item Incorporation of “free” external innovation.
+		\item Motivation: self-sustainable (or less resource-consuming) communities.
+	\end{itemize}
+Two open source examples are the IBM’s WebShere and Apple’s Safari browser.
+Android,KDE .Apache, Konqueror,,Darwin open source project the firm adopting open source components.
+
+ \item \textbf{Donated complements:} firms make their money off of the core innovation, but seek donated labor for valuable complements.
+Firms have indirectly and directly supported users collaboration that is coordinated using open source technologies.
+
+	\begin{itemize}
+		\item Maximization: more value for internal innovation (coreproduct).
+		\item Incorporation: complements are attracted innovation.
+		\item Motivation: developers involved in the core product, but willing more functionality.
+	\end{itemize}
+Examples: Early BSD Unix, Matlab Central, PC Game “Mods”,Avalache .
+\end{itemize}
+
+ %  Amal Roumi
 

### 2013-01-18 MSWL_ECO_lesson4.tex
updated #8 - Summary: Business models Notes from Second part of Business Models in Spanish.

@@
 
- I've commented this section to avoid compiling problems. commet is defined in
- verbarim package, use it as a normal item after importing in main file.
-
-%\begin{comment}
 \subparagraph{Main Content}
@@
 
-\section{Part II}\label{Part II} % TODO: Change for properly section title.
+\section{Where is FLOSS ?}\label{sec:floss-bm}
 
-...
+FLOSS by itself is not, and it has never been, a business model.
+
+\subsection{A guide for SMEs}
+
+Mobiles: sell mobiles as a business model using FLOSS, more viability, cost
+ structure, not the model itself.
+
+If the company
+
+Si la empresa no utilizara FLOSS no sería una empresa viable, es decir, sin el
+ uso de FLOSS la empresa no existiría.
+
+\subsection{FLOSSMETRICS}
+
+Study from 2008/2009 analysing 218 companies that 25\% of their total revenues
+ directly or indirectly from FLOSS.
+
+Identifying common business strategies around FLOSS.
+
+\subsubsection{3 axes}
+
+\begin{itemize}
+    \item Software model
+    \begin{itemize}
+        \item Propietary vs libre software
+    \end{itemize}
+
+    \item Development model
+    \begin{itemize}
+        \item Barries to collaboration
+        \item Single developer/reducted group vs. large communty, global outrech.
+    \end{itemize}
+
+    \item Business model
+    \begin{itemize}
+        \item Type of revenues model.
+        \item Numerous options: Training, support, on-demand changes, productizing, SaaS, etc.
+    \end{itemize}
+\end{itemize}
+
+Development model
+Google Android - Private production, central model.
+GCC - Libre, low barriers, medium group.
+GLibC - Libre, low barriers, little group.
+
+Business model
+SaaS using private model.
+
+\begin{comment}
+\section{Strategic uses of Libre Software}
+
+\emph{No cambias el modelo de negocio tradicional pero te apoyas en productos FLOSS para reforzalo.}
+
+Es muy difícil encontrar una empresa que no utilice ningún producto de FLOSS. FLOSS es adaptativo
+para los requisitos de una empresa. Traducciones, adaptaciones a distintos tipos de hardware,
+interoperabilidad, entornos organizativos, culturales, nuevas tecnologías.
+
+Otra tendencia, es la estrategia de liberar distintas distribuciones de software como FLOSS
+para crear una comunidad libre alrededor.
+
+Apple, Microsoft, Facebook, liberan partes de su código para crear una comunidad para mantener
+el producto por parte de la misma.
+
+Desequilibrios.
+
+Usuarios que no quieren pagar por el Software privativo y se conforman con el programa libre y gratuito.
+Pueden llegar a mejorar al modelo privativo. El privativo se ve desafiado cuando aparece una solución FLOSS
+fiable que genere una aceptación en el mercado. (sitemas embebidos)
+
+\section{Carlo Daffara taxonomy}
+
+\begin{itemize}
+    \item Dual licencing: FLOSS version and propietary version. MySQL Enterprise and MySQL FLOSS.
+    \item Open core: Allows mixing FLOSS and proprietary elements. MSExcange mail model.
+    \item Product specialists: Superior knowledge, additional services. Best service for the product.
+    \item Platform providers: Integration, product testing. PaaS: FLOSS producto integration, glue products. Test by your own.
+    \item Aggregate support providers: Primer nivel de soporte para diferentes tipos de software libre. Escalando el soporte entre todos los niveles.
+    \item Selection/consulting companies: Closer to the analyst role, minimum impact on FLOSS communities.
+    \item Legal certification and consulting : Assessment on license compatibility. Certificate interoperability.
+    \item Training and documentation: Either as part of a broader support contract or companies exclusively devoted to this market area. User guides, documentation, courses.
+    \item R&D cost sharing : Initial investment + creating community to reduce R&D costs.
+Create a community to reduce I+D cost % haciendo que la gente aporte desde fuera para mantenerse al día mediante la herramienta de la comunidad.
+OpenStack from NASA throught RackSpace sharing development cost.
+    \item Indirect revenues: Baseline for sales of associated products or services (commodities).
+\end{itemize}
+
+\subsection{Dual (or multiple) licensing)}
+
+Tienes que ser el dueño de todo el código, los contribuyentes ceden sus derechos de copyright para
+licenciar el producto. El simple hecho de tener que firmar un documento echa hacia atrás las contribuciones de la comunidad.
+
+MySQL integrado en otros productos para que el producto no se liberase siendo privado y de esta forma no violar la licencia GPL.
+Relacionado con la licencia, la licencia Apache y privativo no hubiera funcionado ya que Apache permite el uso de su código dentro
+de un producto privado.
+
+\subsection{Open core}
+
+Los ingresos o beneficios se centran en los componentes privativos alrededor 
+del núcleo libre siendo esa misma empresa la única que puede desarrollar 
+las mejoras de sus servicios. Todo el núcleo es libre.
+
+Weakness: Desarrollar el servicio privativo mediante FLOSS, replicando sus
+funcionalidades, reimplementación de un producto popular.
+
+Suelen tener licencias duales para poder proporcionar la versión FLOSS y la privada.
+
+Base grande de usuario, visibilidad y oportunidad de convertirlos a usuarios premium para 
+obtener los beneficios.
+
+Ganan valor de marca al ser reconocidos por la liberación del core.
+
+Ejemplo: Implementación de Apache SSL. 
+Valor añadido al software al ser la compañía la propietaria del núcleo.
+
+Es un modelo de negocio libre controvertido ya que genera componentes privatidos a 
+través de su plataforma libre, el núcleo.
+
+Hay un compenente de duda alrededor de estos proyectos ya que no se abren a la
+ comunidad al completo al desarrollar recursos privativos para la propia empresa.
+
+Otra forma más positiva de verlo, es que atrae a diferentes empresas para 
+invertir en el producto al poder desarrollar módulos privados. La empresa gasta
+ un gran esfuerzo en la viabilidad del núcleo ya que si no se extiende sus 
+ servicios no tendrían un nicho de negocio.
+
+¿ Gana más recursos ? o se puede ver perjudicada la empresa por este 
+comportamiento del modelo de negocio. ¿ De que depende que le convenga a una 
+empresa adoptar este modelo ?
+
+Es un modelo muy cercano al privativo que intenta de alguna manera mantener al 
+cliente cercado e inmóbil. Este modelo puede tender a desaparecer porque se puede
+dar el caso que otra comunidad, proyecto o empresa que avance más rápido que 
+la solución ofrecida por la empresa y por lo tanto se vea obligada a liberar el 
+código, siendo demasiado tarde.
+
+El modelo se mueve a través de una ventana temporal, es decir el 
+openCore no parece viable de por vida, tiende a la liberación completa 
+del código. Si el producto tiene éxito, es mucho más fácil que ocurra.
+
+SendMail, SugarCRM.
+
+\subsection{Product specialists}
+
+Tengo el mejor conocimiento del producto certificado.
+
+LPI y RedHat: dos certifiaciones alrededor de Linux.
+
+Modelo de Redhat en el campo del could computing: Incluir desarrolladores en el 
+grupo de OpenStack y convertirse en una empresa especialista, no hace falta ser el creador del producto.
+
+Weakness: Que otra empresa con mejor imagen, más nombre, mejores desarrollos desborde tu marca conseguida
+ y por lo tanto quedes relegado a un segundo plano.
+
+\subsection{Platform providers}
+
+\begin{itemize}
+  \item Selection, support, integration and services around a set of projects integrated in a single, tested and verified product.
+  \item Key points; 
+  \begin{itemize}
+    \item Verification services
+    \item Additional services    
+  \end{itemize}
+  \item Copyright ownership prevents direct copy (not cloning).
+  \item Example: Redhat.
+  \item The source code is libre software, but the product name and logo are trademarks.
+  \item Considerable effort to eliminate them from files, doocumentation, etc.
+\end{itemize}
+
+Modelo de negocio de integración: tabletas de marca blanca, software, hardware, etc\ldots
+Reproductores multimedia 'carrefour'.
+
+Especialización y conversión en marca blanca de una tecnología.
+
+\subsection{Aggregate support providers}
+
+Parecido al soporte clásico pero con la diferencia de que es a todos los niveles.
+
+Dos caminos a elegir:
+Mediante el uso de recursos propios o la contratación de terceros para la resolución de los problemas.
+Frontdesk; gestiona las peticiones a través de las empresas que conozcan el producto.
+
+Clear benefit for large projects whose costs could raise due to excessive diversification of support channels (comprehensive help-desl).
+
+OpenLogic: experts in Linux.
+
+\subsection{Selection/consulting companies}
+
+No desarrollan software, se encargan de proveer soluciones asesorando al cliente. Pruebas, análisis de rendimientos, conocimientos del territorio.
+
+Gente experta en un campo dedicado a hacer análisis de un producto en cuestión. 
+Estas empresas proporcionan herramientas que ayudan a reproducir resultados para interpretarlos haciendo análisis más completos.
+
+Open WebApps
+
+\subsection{Legal certification/consulting}
+
+Palamida, Blackduck (mediante ohloh), Sonatype.
+
+Especialización en las licencias de software y análisis de código para buscar posibles incompatibilidades.
+
+\subsection{Training and documentation}
+
+Training and documentation not certifiation. GBDirect.
+
+\subsection{R&D Const sharing}
+
+Memo (nokia), nokia sólo desarrollaría una parte y el resto se quedaría en manos de la comunidad. Este modelo
+ es previo a Android. Sabían del coste de software pero no eran un empresa productora, se dedicaban a la 
+ fabricación de los terminales por lo que optaron por este modelo, dejar que el software creciera de mano
+ de la comunidad. KDE, QT, Linux eran partes que formaban el núcleo de Memo.
+
+OpenStack: RackSpace se dedica al mercado del cloud, vende clouds. Productizó OpenStack compartiendo el gasto 
+del desarrollo con la comunidad (HP, Redhat, Intel\ldots). No es su nicho de mercado y de esta forma el desarollo
+ de software libre lo comparte con los demás para aprovecharse del producto.
+
+\subsection{Indirect revenues}
+
+Intel, Dell, O'Relly. Perl fue el primer caso de la financiación de un proyecto de software libre a Larry Wall 
+para que éste escribiese un libro para su editorial manteniendo los gastos del proyecto a su costa para sacar 
+la próxima versión de Perl.
+
+% Include graphic.
+
+Open source going mainstream - Gartner group report.
+
+Se duplican los beneficios cada tres años después del crecimiento del uso de software libre.
+
+\ldots
+\end{comment}
 

### 2013-01-19 MSWL_ECO_lesson5.tex
Update MSWL_ECO_lesson5.tex
Update MSWL_ECO_lesson5.tex
My Dear classmates .
This is my first latex, Github using  , please let me know if there is something wrong.
Amal
Package inputenc Error: Keyboard character used is undefined

@@
 
-\section{Introduction}
+\section{Introduction}\label{lesson-5-introduction}
 
-\section{Part I}\label{Part I}  % TODO: Change for properly section title.
-
-\section{Part II}\label{Part II} % TODO: Change for properly section title.
-
-% open innovation and open source .
-%Patterns of open innovation in free software
+% open innovation and open source 
 % done by Amal Roumi
 
-\section* {Open source and Open innovation} \label{Open source and Open innovation}
+\section{Open source and Open innovation} \label{Open source and Open innovation}
+\"Open innovation is a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology\"
+'Open innovation' is a term coined by Professor of Business Henry Chesbrough in his 2003 book Open Innovation: The New Imperative for Creating and Profiting from Technology.
 
-“Open innovation is a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology”
-‘Open innovation’ is a term coined by Professor of Business Henry Chesbrough in his 2003 book Open Innovation: The New Imperative for Creating and Profiting from Technology.
+In the years since its publication, Chesbrough\'s ideas on how technology should be managed and exploited have become extremely influential.
+Over the same period, the public profile of free and open source software (FOSS) has risen.
 
-In the years since its publication, Chesbrough’s ideas on how technology should be managed and exploited have become extremely influential.
-Over the same period, the public profile of free and open source software (FOSS) has risen.
 So we can say that FOSS is an example of open innovation in software.
@@
 
-\subsection* {Patterns of open innovation in free software} \label{Patterns of open innovation in free software}
+\subsection{Patterns of open innovation in free software} \label{Patterns of open innovation in free software}
 There are four patterns of open innovation in free software:
@@
  \item  \textbf {Spintouts}: Release the poetical of technologies within the firm that are not creating value.and they have opportunities to release more value from their technologies by position them out the firm but at the same time maintaing an ongoing corporate involvement.
-Here the “spinout”come.
+Here the \"spinout\"come.
 	\begin{itemize}
@@
 	\end{itemize}
-Eclipse, Beehive,Jikes projects the sponsor firms spinout open source project that were closely aligned with firm’s ongoing strategies
+Eclipse, Beehive,Jikes projects the sponsor firms spinout open source project that were closely aligned with firm\'s ongoing strategies
 
@@
 		\item Maximization by centering on core products.
-		\item Incorporation of “free” external innovation.
+		\item Incorporation of "free" external innovation.
 		\item Motivation: self-sustainable (or less resource-consuming) communities.
 	\end{itemize}
-Two open source examples are the IBM’s WebShere and Apple’s Safari browser.
-Android,KDE .Apache, Konqueror,,Darwin open source project the firm adopting open source components.
+Two open source examples are the IBM's WebShere and Apple's Safari browser.
+Android, KDE, Apache, Konqueror, Darwin open source project the firm adopting open source components.
 
@@
 	\end{itemize}
-Examples: Early BSD Unix, Matlab Central, PC Game “Mods”,Avalache .
+Examples: Early BSD Unix, Matlab Central, PC Game "Mods",Avalache .
 \end{itemize}

### 2013-01-19 MSWL_ECO_lesson3.tex
updated #6 - Summary: Cost/benefit analysis
- Section labels.
add list of figures

@@
 
-\section{Introduction}
+\section{Introduction}\label{lesson-3-introduction}
 
@@
 
-\subsection{Cost/benefit analysis}
+\subsection{Cost/benefit analysis}\label{lesson-3-cost-benefit}
 
@@
 
-\section{What disappears when a business uses libre software?}
+\section{What disappears when a business uses libre software?}\label{lesson-3-disappears}
 
@@
             \includegraphics{blackbox}
-            \caption{BlackBox; we only can see the input and the output, not the process.}
-            \label{fig:curve}
+            \caption[BlackBox]{BlackBox; we only can see the input and the output, not the process.}
+            \label{fig:blackbox}
         \end{center}

### 2013-01-20 MSWL_ECO_lesson5.tex
closed #10

@@
 \section{Introduction}\label{lesson-5-introduction}
+FLOSS is an example of model of innovation in software and it is interesting how this model is developed by open communities, in some way similar to another business areas.
+
+Sustainability of a FLOSS project is the capable funding option of being maintained at a steady level, we could identify the following funding:
+ 	\begin{itemize}
+		\item External funding
+		\item Income comes from activities of the organization
+		\item Developments without direct funding
+		\item Developments for internal use
+		\item Mixed models
+ 	\end{itemize}
 
@@
 
+ \section{Sustainability}\label{Sustainability}
+The FLOSS communities and their projects are like living being, therefore they need food to stay alive.
+That is, will the community have the enough resources to be sustainable?, to answer this question is really important for the future of a FLOSS project, and how communities find funds by companies or organizations create different models of sustainability.
+
+ 	\subsection{Funding options for libre software}\label{Funding options for libre software}
+ 	Initially this classification was treated for business model.
+
+		\subsubsection{External funding}
+
+		Public administrations usually have the following motivations:
+			\begin{itemize}		
+				\item \emph{Scientific motivation}: that software required to produce certain results must be available, in order to reproduce them.
+				\item \emph{Precompetitive motivation}: all industry network can benefit from precompetitive results. 
+				\item \emph{Promoting standards motivation}: Implementing reference versions of a certain standard.
+
+				\item \emph{Social motivation}: Funding creation of common basic infrastructure of information society.
+
+				\item \emph{Case study}: Gnat (Ada compiler), \$1M assigned from U.S. Government to NYU for its development. And the curious was the development of this product was initially valued around \$100M.
+			\end{itemize}
+
+			For instance, the German Government has issued a DM 250,000 grant to the GNU Privacy Guard (GPG)~\cite{GPG} project.
+The grant will help the GPG team develop for more platforms and develop easy to use interfaces and APIs.
+Another examples of funding by public administrations are the Spanish States Government,such as Extremadura, Andalucia, Valencia and others.
+
+			Others external fundings:
+
+			\begin{itemize}	
+			\item \emph{Non-profit organization}: This foundation usually made for non-profit orgarnization.
+The funding mechanisms can be direct or indirect.
+For example, \emph{MediaWiki} is a free wiki software application, developed by the Wikimedia Foundation and others, it is used to run all of the projects hosted by the Foundation, including Wikipedia, Wiktionary and Commons.
+\emph{Blender} is a free and open-source 3D computer graphics software product used for creating animated films, visual effects, interactive 3D applications or video games; Blender was developed as an in-house application by the Dutch animation studio Neo Geo and Not a Number Technologies (NaN).
+The \emph{Open Bioinformatics Foundation} is a non profit, volunteer run organization focused on supporting open source programming in bioinformatics.
+			\item \emph{Indirect}: Some companies funds FLOSS projects to achive profits with something associated with this project.
+Therefore Google funds Android project because it wants to achieve mobile phone technology positioning.
+Another example is O'Reilly that funds FLOSS projects to sell books about them.
+Many hardware companies funds FLOSS projects to improve their firewires and drivers.
+Red Hat, a company of services, funds Gnome projects because its core product uses it.
+
+			\item \emph{Need of improvements}: Some companies funds FLOSS project to improve requirements of their products, saving resources in contract with the traditional development process.
+So for example Corel funds Macadamian to Wine to achieve a Corel Draw working in linux.
+
+			\end{itemize}
+
+		\subsubsection{Income comes from activities of the organization(s)}
+			\begin{itemize}	
+				\item \emph{Detailed knowledge}: Corporations fund in some leading project, usually for having detailed knowledge about it.
+For example, many corporations are collaborating in WebKit project to achieve in mobile phone market, besides small ones like Igalia can achieve this technology positioning, and in with detailed knowledge they can sell its services.
+
+Autofinanciado: venta de marca
+%-----------------------------------
+				\item \emph{Brand selling}: If a brand is sufficiently known, it is easier to sell services around it.
+For examble, Mozilla with Firefox is using this model.
+Even more clear, it is Red Hat with its core product (Enterprise Red Hat OS), and so brand "Red Hat" means detailed knowledge of linux for customers.
+
+			\end{itemize}
+		\subsubsection{Developments without direct funding} 
+		Most of the projects of FLOSS are development in this way.
+In many cases, there are indirect funding: Some companies dedicate human resources to develop the project for different reasons, some organization fund this project to improve some feature interested for them, donation of money or IT infrastructure.
+		For example, initially OpenStack is here, after it had a relative success, a foundation was created with \$10M to promote this project.
+
+		\subsubsection{Developments for internal use}
+		For instance, Cisco used this way of funding for development of tools designed to optimize the management of large numbers of printers (CEPS, Cisco Enterprise Print System ~\cite{CEPS}).
+
+		\subsubsection{Mixed models}
+		The reality is that the models are mixed and project usually move among them along its history.
+
 \section{Conclusions}\label{conclusions}
+Open innovation is "new" paradigm, which is working properly in FLOSS project, but it is more important to understand that it is a model that can work perfect many areas.
+Sustainability is essecial feature for FLOSS project viability, there are many way for fundings, and projects and organizations usually use mixed models and change along the life of the project.

### 2013-01-20 MSWL_ECO_lesson4.tex

--- 2012-12-28 MSWL_ECO_lesson4.tex
@@
 
-\section{Part II}\label{Part II} % TODO: Change for properly section title.
+\section{A guide for SMEs}
 
-...
+FLOSS by itself is not, and it has never been, a business model.
+
+\subsection{Where is FLOSS ?}\label{sec:floss-bm}
+
+Mobiles: sell mobiles as a business model using FLOSS, more viability, cost
+ structure, not the model itself.
+
+If the company
+
+Si la empresa no utilizara FLOSS no sería una empresa viable, es decir, sin el
+ uso de FLOSS la empresa no existiría.
+
+\subsection{FLOSSMETRICS}
+
+Study from 2008/2009 analysing 218 companies that 25\% of their total revenues
+ directly or indirectly from FLOSS\footnote{http://libresoft.es/research/projects/flossmetrics}.
+
+FLOSSMetrics stands for Free/Libre Open Source Software Metrics. 
+The main objective of FLOSSMETRICS was to construct, publish and analyse a large
+ scale database with information and metrics about libre software development 
+ coming from several thousands of software projects, using existing 
+ methodologies, and tools already developed.
+
+\subsubsection{3 axes}
+
+\begin{itemize}
+    \item Software model
+    \begin{itemize}
+        \item Propietary vs libre software
+    \end{itemize}
+
+    \item Development model
+    \begin{itemize}
+        \item Barries to collaboration
+        \item Single developer/reducted group vs. large communty, global outrech.
+    \end{itemize}
+
+    \item Business model
+    \begin{itemize}
+        \item Type of revenues model.
+        \item Numerous options: Training, support, on-demand changes, productizing, SaaS, etc.
+    \end{itemize}
+\end{itemize}
+
+Development model samples:
+\begin{itemize}
+  \item Google Android - Private production, central model.
+  \item GCC - Libre, low barriers, medium group.
+  \item GLibC - Libre, low barriers, little group.
+\end{itemize}
+
+\section{Strategic uses of Libre Software}
+
+\emph{You do not change the traditional business model, but you lean on FLOSS
+ products to enhance the model.}
+
+It is very difficult to find a company that does not use any of FLOSS. 
+FLOSS is adaptive to the requirements of a company. 
+Translations, adaptations to different types of hardware, interoperability, 
+organizational environments, culture, new technologies.
+
+Otra tendencia, es la estrategia de liberar distintas distribuciones de 
+software como FLOSS para crear una comunidad libre alrededor.
+
+Apple, Microsoft, Facebook, release parts of your code to create a community to 
+keep the product by the same community (composed of these same companies also).
+
+\subsection{Imbalances}
+
+Users who do not want to pay for proprietary software and are satisfied with 
+the program free of charge.
+
+They can improve the proprietary model. The proprietary is challenged when a 
+FLOSS reliable solution generates an acceptance in the market. (sitemas embedded)
+
+\section{Carlo Daffara taxonomy}
+
+\begin{itemize}
+    \item Dual licencing: FLOSS version and propietary version. 
+    MySQL Enterprise and MySQL FLOSS.
+    \item Open core: Allows mixing FLOSS and proprietary elements. 
+    MSExcange mail model.
+    \item Product specialists: Superior knowledge, additional services. 
+    Best service for the product.
+    \item Platform providers: Integration, product testing. 
+    PaaS: FLOSS producto integration, glue products. Test by your own.
+    \item Aggregate support providers: 
+    First level of support for different types of free software. 
+    Scale the support from all levels.
+    \item Selection/consulting companies: Closer to the analyst role, 
+    minimum impact on FLOSS communities.
+    \item Legal certification and consulting : Assessment on license 
+    compatibility. Certificate interoperability.
+    \item Training and documentation: Either as part of a broader support 
+    contract or companies exclusively devoted to this market area. U
+    ser guides, documentation, courses.
+    \item R\&D cost sharing : Initial investment + creating community to 
+    reduce R\&D costs. Create a community to reduce I+D cost
+    Making people contribute from outside to keep up with the community tool.
+    OpenStack from NASA throught RackSpace sharing development cost.
+    \item Indirect revenues: Baseline for sales of associated products or 
+    services (commodities).
+\end{itemize}
+
+\subsection{Dual (or multiple) licensing)}
+
+You must be the owner of all code. Contributors transfer their copyright to
+license the product. The simple fact of having to sign a document pulls back 
+the community contributions.
+
+Example of MySQL.
+Integrated with other products to the product not being liberate private and 
+thus not violating the GPL.
+Related to the license, the Apache License and proprietary, had not worked 
+since Apache allows the use of their code in a private product.
+
+\subsection{Open core}
+
+Revenues or profits focus on proprietary components around
+Open Core. Being the same company, the only one that can develop
+improvement of its services. The entire core is free.
+
+Weakness: Develop proprietary service by FLOSS, replicating their
+functionalities, reimplementation of a popular product.
+
+They usually have dual licenses to provide private and FLOSS version.
+
+Large user base, visibility and opportunity to convert to premium users to get 
+the benefits.
+
+Win brand value to be recognized by the release of the core.
+
+Example: Implementing SSL Apache. 
+Value added to the software to be the company's proprietary core.
+
+It's a business model that generates free controversial and private components
+ through its open platform, the core.
+
+There is a doubt compenente around these projects, as they are not open to the
+  whole community, to develop proprietary resources for the company.
+ 
+A more positive view is that attracts different companies for
+invest in the product to be able to develop private modules. The company spends
+  a great effort on the viability of the core because if not extended their
+  services would not have a business niche.
+
+Do you earn more resources? or may be impaired by this company
+behavior of the business model. What determines which one suits you
+companies adopt this model?
+
+It is very close to the proprietary model, and somehow try to keep the customer 
+fenced and locked. This model may tend to disappear because it may be the case 
+that another community project or company to advance faster than the solution 
+offered by the company and therefore is forced to release the code, being
+ too late.
+
+The pattern moves across a time window, ie not seem feasible OpenCore lifetime 
+tends to complete release code. If the product is successful, 
+it is easier to occur.
+
+Samples: SendMail, SugarCRM.
+
+\subsection{Product specialists}
+
+\emph{I have the best knowledge of the certified product.}
+
+LPI y RedHat: two certifications of Linux knowledge.
+
+Redhat model in the field of computing Could: Include developers in
+OpenStack group and become a specialist. No need to be the creator of the 
+product.
+
+Weakness: another company with better image, more name, best developments, 
+overflowing your brand achieved and therefore, 
+you become relegate to the background.
+
+\subsection{Platform providers}
+
+\begin{itemize}
+  \item Selection, support, integration and services around a set of projects 
+  integrated in a single, tested and verified product.
+  \item Key points; 
+  \begin{itemize}
+    \item Verification services
+    \item Additional services    
+  \end{itemize}
+  \item Copyright ownership prevents direct copy (not cloning).
+  \item Example: Redhat.
+  \item The source code is libre software, but the product name and logo 
+  are trademarks.
+  \item Considerable effort to eliminate them from files, doocumentation, etc.
+\end{itemize}
+
+Integrated business model: private label tablets, software, hardware, etc.\ldots
+Multimedia players from 'Carrefour'.
+
+Specialization and white label conversion technology.
+
+\subsection{Aggregate support providers}
+
+Similar to classic support but with the difference that is at all levels.
+
+Two paths to choose from:
+\begin{itemize}
+  \item By using its own resources or outsourcing to solve problems.
+  \item Frontdesk; handles requests by companies who know the product.
+\end{itemize}
+
+Clear benefit for large projects whose costs could raise due to excessive
+ diversification of support channels (comprehensive help-desl).
+
+OpenLogic: experts in Linux.
+
+\subsection{Selection/consulting companies}
+
+No software developers. They are responsible for advising the client to 
+provide solutions. Testing, performance analysis, knowledge of the territory.
+
+People expert in a field dedicated to doing analysis of a product.
+
+These companies provide tools that help to interpret results by playing 
+more complete analysis.
+
+Open WebApps
+
+\subsection{Legal certification/consulting}
+
+Specializing in software licensing and code analysis for possible 
+incompatibilities.
+
+Palamida, Blackduck (mediante ohloh), Sonatype.
+
+\subsection{Training and documentation}
+
+Training and documentation not certifiation.
+
+GBDirect.
+
+\subsection{R\&D Const sharing}
+
+\emph{Memo (nokia)}, nokia develop only a portion and the rest would be up to the
+ community. This model is before Android. They knew the cost of software but 
+ were not a production company, engaged in the manufacture of terminals so 
+ opted for this model, let the software grew hand in the community. KDE, QT, 
+ Linux were parties that formed the core of Memo.
+
+\emph{OpenStack}: Rackspace is dedicated to the cloud market, sold clouds. 
+RackSpace produced 'OpenStack' sharing the cost of developing the community 
+(HP, Redhat, Intel\ldots). Not your niche market, so the development of free
+ software is shared with others to take advantage of the product.
+
+\subsection{Indirect revenues}
+
+Intel, Dell, O'Reilly. Perl was the first case of financing a free software
+ project to Larry Wall. For Larry Wall wrote a book for his publisher (O'Reilly)
+  paying the cost of the project at its own expense to get 
+  the next version of Perl.
+
+Open source going mainstream - Gartner group 
+report\footnote{http://www.gartner.com/newsroom/id/593207}
+
+\emph{By 2012, 80 per cent of all commercial software will include elements of
+ open-source technology. Many open-source technologies are mature, stable and
+  well supported. They provide significant opportunities for vendors and users
+   to lower their total cost of ownership and increase returns on investment. 
+   Ignoring this will put companies at a serious competitive disadvantage. 
+   Embedded open source strategies will become the minimal level of investment
+    that most large software vendors will find necessary to maintain
+     competitive advantages during the next five years.}
 

### 2013-06-04 MSWL_ECO_lesson1.tex
Delete section part II in lesson 1
Revision chapter 1

@@
 Apache has reached to be a standard de facto, so to sell a Web Server different from Apache, you can not sell anything the same or with less performance than Apache, especially when it is free.\\
+\begin{figure}[htb]
+	\begin{center}
+		\includegraphics{webServer1303}
+		\caption{Web Server share market}
+		\label{fig:webServer1303}
+	\end{center}
+\end{figure}
 \\
@@
 Curious as Google and Apple fight each other in many other businesses.\\
+\begin{figure}[htb]
+	\begin{center}
+		\includegraphics{mobile_tablet_browser}
+		\caption{Mobile-Tablet browser}
+		\label{fig:mobile_tablet_browser}
+	\end{center}
+\end{figure}
 \\
@@
 \\
+\begin{figure}[htb]
+	\begin{center}
+		\includegraphics{Operating_systems_used_on_top_500_supercomputers}
+		\caption{Operatingvsystems used on top 500 supercomputers}
+		\label{fig:Operating_systems_used_on_top_500_supercomputers}
+	\end{center}
+\end{figure}
 To summarize, it has to be stated that FLOSS is a particular case, in technology and, of course, in economics.\\\\
@@
    \hline
-   Decentralized  & Price Based    & \textbf{Social Sharing/Exchange} \\
+   Decentralized  & Price Based    & \textbf{Social Sharing/Exchange} *\\
    \hline
@@
 \\
-(*1)In past time this paradigm was used for small projects (look after of a garden in a neighbourhood)\\
+(*)In past time this paradigm was used for small projects (look after of a garden in a neighbourhood)\\
 It is normally valid for "non-material" things, changing the sense of property. That happens on software, music, etc.\\
@@
 
-\section{Part II}\label{Part II} % TODO: Change for properly section title.
-
 \section{Conclusions}\label{conclusions}
+The evidence from the study of the market numbers suggests that the number of FLOSS projects has grown exponentially at least years, so FLOSS market is not an insignificant market.
+There are niches where the lead product are free software and it is really stable market, and is difficult to change for privative software.
+\\
+Nowadays, free software is sometimes the best option in certain business and sometimes not, but the important thing to note is that almost all companies are considering free software strategies on their global business plans.
+\\
+It is important to consider that internet is changing the model.
+Many actors now have the tool to be interconected and is used to synchronise people in a very efficient way.
+A new paradigm in internet springs up strongly upwards.
+Around this new concepts appear like crowdsourcing, which is an online, distributed requirement solving and production mode, and not only for software production.

### 2013-06-08 MSWL_ECO_lesson3.tex
Fix position of black box image
Revision chapter 3

@@
         \item Compare benchmarks between projects and its goals.
-Could analyse every code line as you want and check every feature that has the product and compare with other projects using standar quality models\footnote{QSOS, OpenBRR, QUALOSS} adapted to your interests.
+Could analyse every code line as you want and check every feature that has the product and compare with other projects using standard quality models\footnote{QSOS, OpenBRR, QUALOSS} adapted to your interests.
 And of course, anyone can view the source code of your project so you have to get it right and well documentated.
@@
 Maintenance is very important and every new business model could grow around FLOSS projects.
-So your product as a costumer depends only of one company, what would happen if the company goes bankrupt?
+So your product as a costumer depends only on one company, what would happen if the company goes bankrupt?
     \end{itemize}
@@
 This application hides we treat our data to drive the process of sending an email and the result shows that message has been sent as output
-    \end{itemize}  
-    \begin{figure}[h]
+            \begin{figure}[htp]
                 \begin{center}
@@
             \end{figure}
+
     \end{itemize}
+
+\end{itemize}
+
+\pagebreak
 
@@
 
-\textbf{TBC}
+Companies and public institutions would base its strategy on FLOSS because libre software is a viable business option, with high quality products, avoiding lock vendor, and technical reliability;  allows for new business opportunities, reducing barriers, incresing collaboration, and Libre software as an strategic tool; but it is important to know how to take advantage of those.
+
+When in a business segment actors start to use free software, preventing in this niche monopolies or oligopolies and black boxes products, creating reliability products and available to test, and preventing dependency of strategies of providers.

### 2013-06-08 MSWL_ECO_lesson4.tex
Revision chapter 4

@@
 
-FLOSS
-
-The one million dollar question \emph{Which model has success.}
+The one million dollar question \emph{Which model has success?.}
 
@@
 
-If the company
-
-Si la empresa no utilizara FLOSS no sería una empresa viable, es decir, sin el
- uso de FLOSS la empresa no existiría.
+If the company did not use FLOSS would not be a viable, that is,  without the
+  FLOSS use the company would not exist.
 
@@
 
-Otra tendencia, es la estrategia de liberar distintas distribuciones de 
-software como FLOSS para crear una comunidad libre alrededor.
+Another trend is the strategy of releasing different distributions of
+FLOSS software to create a free community around.
 
@@
 \section{Conclusions}\label{conclusions}
+
+The business model must always pivot on the use value, not on the sale value.
+\\
+
+There are 3 axes influencing  software landscape: Software model, development model, and  business model.
+
+Nowadays, use of libre software would reach more than 90 \% of companies worldwide. We could say that almost every company has a strategy about libre software.
+
+The  Carlo Daffara taxonomy (FLOSSMETRICS) of business model and examples
+\begin{itemize}
+    \item         Dual licensing, e.g. MySQL, Mozilla software, Sleepycatdb, Pentaho
+    \item         Open Core, e.g. Zimbra, SugarCRM, Eucalyptus, SendMail
+     \item        Product specialists, e.g. Alfresco
+    \item        Platform Providers, e.g. RedHat, OpenStack
+     \item        Selection/consulting companies, e.g. Navica, Open WebApps
+     \item        Aggregate support providers, e.g. OpenLogic
+     \item        Legal certification and insurance, e.g. Palamida
+    \item         Training and documentation, e.g. Gbdirect
+     \item        R\&D cost sharing, e.g. Eclipse, Webkit, OpenStack, Maemo (Nokia)
+     \item        Indirect revenues, e.g. Firefox, Android, Perl (O'Reilly)
+\end{itemize}
